---
title: "Inference SPR results: Preprocessing"
author: "Miriam Schulz"
date: "9/15/2022"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# About

This R markdown script preprocesses the raw SPR results from PCIbex using custom functions stored in `functions.R`.

Reads in the raw `results.csv` file downloaded from PCIbex and merges it with the items + pretest results.

Returns two dataframes:

- `results_reads.csv` contains the reading time and reaction time data 
- `results_survey.csv` contains the demographics and survey data

Summary of data preprocessing steps: 

- exclude participant with low accuracy and reading times
  (pre-determined)
- add unique subject numbers (1-40)
- annotate list order, list number and trial number
- annotate word position with respect to the critical target word 
  (target = 0, precritical = -1, spillover1 = 1, etc)
- annotate plausibility rating accuracy
- merge data with the items & pretest result file
- export data


# Preliminaries

```{r, include = FALSE}
# Workspace, libraries, functions
rm(list = ls())  # clear workspace (optional)
library(tidyverse)
source("functions.R") # custom functions 
```

Read in data, remove participants and add Subject IDs:

```{r collapse=TRUE}
# Read in raw data
df <- readPCibex("results.csv")

# Remove participants
# Participant with very long RTs who timed out: 
exclude.participants <- c("42540fa1be6151d9f8ae49bb3098494c")
df <- filter(df, !IPhash %in% exclude.participants)

# Add subject numbers:
df.subjects <- data.frame(IPhash = unique(df$IPhash),
                          Subject = 1:length(unique(df$IPhash)))
# use plyr::join instead of merge to preserve original row ordering:
df <- plyr::join(df.subjects, df)
```

Next, split the data into separate data frames for reading times and survey/demographics.

The `getReads()` function annotates:

- list order, Latin list number and trial position
- word position with respect to the critical target word 
- plausibility ratings: annotate response accuracy for each trial

```{r collapse=TRUE}
reads <- getReads(df)
survey <- getSurvey(df)
```


# Add pretest results

Read in the items + pretest results and merge with the reading time data frame:

```{r}
# Open items file with pretest results
items <- read.csv("items_pretests.csv", header = TRUE)

# Add new item numbers:
items$ItemPretests <- items$Item  # save original item number
items$Item <- rep(1:40, each=4)

# Select columns 
items <- items[, c(1,20,2:6)]

# Merge and reorder
reads.export <- merge(items, reads, by=c("Item", "Cond"), all.y = TRUE)
reads.export <- reads.export[, c(1, 3, 2, 4:11, 13, 27, 22, 15:19, 23:26)]
reads.export <- reads.export[ with(reads.export, order(Item, Cond, Region)), ]
```

# Export 

Two separate files are exported:

1. `results_reads.csv` for the reading time data, and
2. `results_survey.csv` for the survey/demographics data, including age, gender, handedness, language background and post-experimental survey data

```{r}
write.csv(reads.export, "results_reads.csv", row.names = FALSE)
write.csv(survey, "results_survey.csv", row.names = FALSE)
```

