---
title: "SPR results: Plausibility ratings"
author: "Miriam Schulz"
date: "9/15/2022"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, fig.width=6, fig.height=5)
```

# About

In this script:

Analysis of the plausibility ratings data (the online task during the Self-Paced Reading experiment), split into:

- Rating accuracy
- Rating reaction times


# Preprocessing

```{r include = FALSE}
# Workspace, libraries, functions
rm(list = ls())       # clear workspace (optional)
library(tidyverse)
library(MASS)         # for boxcox function 
library(grid)         # for textGrob()
library(gridExtra)    # for grid.arrange
source("functions.R") # custom functions 
```

Read in data and filter to a single region for the reaction time analysis:

```{r}
df <- read.csv("results_reads.csv")
ratings <- removeRegions(df, keep.regions=0)
```

The range of the reaction times above shows that no extreme outliers need to be removed before proceeding.

With respect to the accuracy, this script revealed that three items have low accuracy in the B condition, suggesting that the inference is weaker in these items, even though they passed the criteria for the pretests.
In order to run this script without these items and compare results, uncomment the following line:

```{r}
#ratings <- filter(ratings, !Item %in% c(5, 13, 25))
```


# Rating accuracy

## Grand mean and sd

```{r}
meanRound(ratings$RatingCorrect, 2)
meanRound(sd(ratings$RatingCorrect))
```

## Accuracy by subject

```{r}
subj.accs <- aggregMeans(ratings, as.formula(RatingCorrect ~ Subject),
                         round.data=TRUE)

# All subject accuracies:
subj.accs

# Range, mean, SD:
round(range(subj.accs$Mean), 2)
meanRound(subj.accs$Mean)
meanRound(sd(subj.accs$Mean))
```

## Accuracy by condition

```{r}
cond.accs <- aggregMeans(ratings, as.formula(RatingCorrect ~ Cond), 
                         round.data=TRUE)

# By-condition accuracy and SD:
cond.accs

# Range:
round(range(cond.accs$Mean), 2)
```

## Accuracy by item

Check if some particular items or fillers had low overall accuracy ($< 0.7$)

```{r}
item.accs <- aggregMeans(ratings, as.formula(RatingCorrect ~ Item), 
                         round.data=TRUE)
range(item.accs$Mean)
# Print problematic items with low overall accuracy:
filter(item.accs, Mean < 0.7)
# Range of overall accuracy for the experimental items only (no fillers):
range(filter(item.accs, Item < 100)$Mean)
```

**Note:** Low mean accuracy items (mean accuracy $< 0.7$) are all fillers, no experimental item is concerned.

## Accuracy by item + condition (experimental items only)

```{r}
ratings.items <- removeFillersPractice(ratings)
item.cond.accs <- aggregMeans(ratings.items,
                              as.formula(RatingCorrect ~ Item + Cond), 
                         round.data=TRUE)
range(item.cond.accs$Mean)

item.cond.accs <- dplyr::select(item.cond.accs, Cond, Item, Mean)
item.cond.accs <- item.cond.accs %>% 
  spread(Cond, Mean)
item.cond.accs <- merge(item.cond.accs, item.accs)
item.cond.accs
item.cond.accs <- item.cond.accs[, 1:6]

# Print any items that had accuracy < 70% for any individual condition:
item.cond.accs %>% filter_all(any_vars(. < 0.7))
```

**Note:** 3 experimental items have low accuracy in one condition (always in Condition B).

## Plot: Accuracy by condition

```{r echo=FALSE, fig.width=5, fig.height=4}
p <- cond.accs %>% 
  ggplot(aes(x = Cond, y = Mean, fill=Cond)) + 
  geom_bar(stat = "identity",
           width=0.8) +
  geom_errorbar(
    aes(ymin = Mean - SE, ymax = Mean + SE),
    width = 0.15,
    linewidth = 0.3
  ) +
  xlab("Condition") +
  ylab("Accuracy (%)") +
  #ggtitle("Mean rating accuracy per condition (error bars = 1 SE)") +
  scale_fill_manual(values = c("cornflowerblue", "chartreuse3",
                                "tomato2", "darkgoldenrod1",
                                "gray85","gray75")) +
  scale_x_discrete(labels=c("A", "B", "C", "D",
                            "Plausible\nfillers", "Implausible\nfillers")) +
  theme_minimal() + 
  theme(legend.position = "none",
        text = element_text(size = 13))
filename <- paste0("./plots/", "rating_accs.pdf")
ggsave(filename, plot = p, width = 5, height = 4)
p + ggtitle("Mean rating accuracy per condition")
```


# Rating reaction times

## Mean and SD before outlier removal

```{r}
# Decision Reaction Times
round(mean(ratings$DecisionRT))
round(sd(ratings$DecisionRT))
range(ratings$DecisionRT)
```


## Remove outliers 

First, inspect the histograms and QQ plots of the reaction times to visually check for outliers:

```{r echo=FALSE, fig.width=10, fig.height=5}
par(mfrow=c(1,2))

# Histogram of RTs:
hist(ratings$DecisionRT,
     breaks = 50,
     col='steelblue',
     main='Raw rating decision times')

# Histogram of log RTs:
hist(log(ratings$DecisionRT),
     breaks = 50,
     col='steelblue',
     main='Log-transformed rating decision times')

# QQ plots

# Untransformed RTs
qqnorm(ratings$DecisionRT,
       col='steelblue',
       pch=16,
       main='QQ plot of raw decision times')

# Log-transformed RTs
qqnorm(log(ratings$DecisionRT),
       col='steelblue',
       pch=16,
       main='QQ plot of log-transformed decision times')
par(mfrow=c(1,1))  # reset
```

The QQ plots show the presence of outliers. 
Therefore, rating reaction times were filtered to lie within $\pm 2.5$ standard deviations from the by-participant mean, resulting in the loss of $3.34\%$ of the data:

```{r}
ratings <- removeOutliersRatings(ratings,
                                 method="sd",
                                 sd=2.5)
```

Check the plots again after removing outliers:

```{r echo=FALSE, fig.width=10, fig.height=5}
par(mfrow=c(1,2))
# Histogram of RTs:
hist(ratings$DecisionRT,
     breaks = 50,
     col='steelblue',
     main='Raw rating decision times')

# Histogram of log RTs:
hist(log(ratings$DecisionRT),
     breaks = 50,
     col='steelblue',
     main='Log-transformed rating decision times')

# QQ plots
# Untransformed RTs
qqnorm(ratings$DecisionRT,
       col='steelblue',
       pch=16,
       main='QQ plot of raw decision times')

# Log-transformed RTs
qqnorm(log(ratings$DecisionRT),
       col='steelblue',
       pch=16,
       main='QQ plot of log-transformed decision times')
par(mfrow=c(1,1))  # reset
```

## Mean and SD after outlier removal

```{r}
# Decision Reaction Times
round(mean(ratings$DecisionRT))
round(sd(ratings$DecisionRT))
range(ratings$DecisionRT)
```

## Reaction time by condition

```{r}
# Mean and SE of ratings, outliers removed
rating.cond.means <- aggregMeans(ratings,
                                 as.formula(DecisionRT ~ Cond)) %>%
  mutate_if(is.numeric, round)
rating.cond.means
```

## Plot: Reaction times by condition

```{r echo=FALSE, fig.width=5, fig.height=4}
# As barplot:
p <- rating.cond.means %>% 
  ggplot(aes(x = Cond, y = Mean, fill = Cond)) + 
  geom_bar(stat = "identity",
           width=0.8) +
  geom_errorbar(
    aes(ymin = Mean - SE, ymax = Mean + SE),
    width = 0.15,
    linewidth = 0.3
  ) +
  xlab("Condition") +
  ylab("Reaction time (ms)") +
  scale_fill_manual(values = c("cornflowerblue", "chartreuse3",
                                "tomato2", "darkgoldenrod1",
                                "gray85","gray75")) +
  scale_x_discrete(labels=c("A", "B", "C", "D",
                            "Plausible\nfillers", "Implausible\nfillers")) +
  theme_minimal() + 
  theme(legend.position = "none",
        text = element_text(size = 13))
filename <- paste0("./plots/", "rating_RT.pdf")
ggsave(filename, plot = p, width = 5, height = 4)
p + ggtitle("Mean rating RT per condition")
```


```{r include=FALSE, fig.width=10, fig.height=4}
#As boxplots (show outliers):

p1 <- ratings %>% 
  ggplot(aes(x = Cond, y = DecisionRT, fill=Cond)) + 
  geom_boxplot() + 
  xlab("Condition") +
  ylab("Reaction time (ms)") +
  scale_fill_manual(values = c("cornflowerblue", "chartreuse3",
                                "tomato2", "darkgoldenrod1",
                                "gray85","gray75")) +
  scale_x_discrete(labels=c("A", "B", "C", "D",
                            "Plausible\nfillers", "Implausible\nfillers")) +
  theme_minimal() +
  theme(legend.position="None",
        text = element_text(size = 13))
filename <- paste0("./plots/", "rating_RT_boxplot.pdf")
ggsave(filename, plot = p1, width = 5, height = 4)
#p1 + ggtitle("Rating reaction time by condition")

p2 <- ratings %>% 
  ggplot(aes(x = Cond, y = log(DecisionRT), fill=Cond)) + 
  geom_boxplot() + 
  xlab("Condition") +
  ylab("Reaction time (ms)") +
  scale_fill_manual(values = c("cornflowerblue", "chartreuse3",
                                "tomato2", "darkgoldenrod1",
                                "gray85","gray75")) +
  scale_x_discrete(labels=c("A", "B", "C", "D",
                            "Plausible\nfillers", "Implausible\nfillers")) +
  theme_minimal() +
  theme(legend.position="None",
        text = element_text(size = 13))
filename <- paste0("./plots/", "rating_RT_boxplot_log.pdf")
ggsave(filename, plot = p2, width = 5, height = 4)
#p2 + ggtitle("Rating reaction time by condition (log-transformed)")

grid.arrange(p1, p2,
             ncol=2,
             top = textGrob("Reaction times by condition (left: untransformed, right: log-transformed)",
             gp=gpar(fontsize=18, fontface=2)))
```
