---
title: "rRTs (script version with reduced number of models)"
author: "Miriam Schulz"
date: "30/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.height = 6, fig.width = 9)
```


## About

This script contains an rRT analysis of the inference reading time data (SPR1) based on the rERP approach in Brouwer, Delogu & Crocker (2020). 


## Read data and preprocess

Read in the data, scale the predictors, etc.

```{r include = FALSE}
# Workspace and libraries
rm(list = ls())
library(tidyverse)
library(lme4)
library(broom)      # for tidy(modeloutput)
library(gridExtra)  # for grid.arrange()
library(grid)       # for textGrob() => change grid.arrange plot name font size 
library(sciplot)    # for se() 
library(RColorBrewer)
source("results_SPR_functions.R")      # preprocessing and helper functions
source("results_SPR_functions_rRT.R")  # rRT functions
```


```{r}
# Read data
df <- read.csv("results_reads.csv", header = TRUE)

# Exclude filler and practice trials
df <- filter(df, Item < 100)

# Transform variables to factors
df$Cond <- as.factor(df$Cond)
df$Item <- as.factor(as.numeric(as.character(df$Item)))
df$Subject <- as.factor(as.numeric(as.character(df$Subject)))

# Transform the binary plausibility predictor to numeric, then to factor
df$Plausible <- as.factor(ifelse(df$Plausible == "Yes", 1, 0))

# Exclude data
df <- filterRTs(df, criteria = c(50, 2500), method = "thresholds",
                remove.incorrect = TRUE, remove.contexts = 500,
                regions = -2:2)

# Extract precritical RT -1 and add to df:
df.precrit1 <- filter(df, Region == -1) %>% 
  dplyr::select(Item, Cond, Subject, RT) %>% 
  rename(RT_precrit1 = RT)
df <- merge(df, df.precrit1, by = c("Item", "Cond", "Subject"))

# Extract precritical RT -2 and add to df:
df.precrit2 <- filter(df, Region == -2) %>% 
  dplyr::select(Item, Cond, Subject, RT) %>% 
  rename(RT_precrit2 = RT)
df <- merge(df, df.precrit2, by = c("Item", "Cond", "Subject"))

# Add log-transformed RTs
df$logRT <- log(df$RT)
df$logRT_precrit1 <- log(df$RT_precrit1)
df$logRT_precrit2 <- log(df$RT_precrit2)

# Rename variables 
rating_colnames <- str_remove(names(df), "Rating")
names(df) <- rating_colnames
df <- rename(df, ProdNorms = ProductionPercentage)

# Order df by Region
#df <- arrange(df, Region, Item, Cond)
df <- arrange(df, Region)

# Scale predictors (careful: needs to be performed AFTER removing data)
df$Association <- scale(df$Association)
df$Inference <- scale(df$Inference)
df$Coherence <- scale(df$Coherence)
df[, c("RT_precrit1", "RT_precrit2","logRT_precrit1","logRT_precrit2")] <- 
  lapply(df[, c("RT_precrit1", "RT_precrit2","logRT_precrit1","logRT_precrit2")], scale)
```


### Define model parameters

Define general model parameters:

- regions to model
- whether to use raw or log-transformed RTs
- which kind of model to run:
  - "lm": simple LM
  - "lm.by.subj": by-subject LMs
  - "lmer": LMER

```{r}
regions <- -1:2
log.RTs <- TRUE
m.type = "lm.by.subj"

if (log.RTs == TRUE) {
  DV <- "logRT"
  precrit1 <- "logRT_precrit1"
  precrit2 <- "logRT_precrit2"
  y.unit <- "Log RT"
  y.range <- c(5.55, 5.70)
  y.range.res <- c(-0.07, 0.07)
} else {
  DV <- "RT"
  precrit1 <- "RT_precrit1"
  precrit2 <- "RT_precrit2"
  y.unit <- "Raw RT (ms)"
  y.range <- c(260, 320)
  y.range.res <- c(-20, 20)
}
```


### Inspect the spread of variation across regions 

Check the *range* of the reading times in the critical regions (i.e. the spread of the variation).

Larger ranges implies that the residuals will be larger due to the increased amount of gvariation in that region.


```{r}
# Density plot
ggplot(filter(df, Region %in% -2:2),
       #aes(x = RT, group = Region, fill = factor(Region))) +
       aes(x = RT, group = Region, color = factor(Region))) +
  geom_density(adjust=0.8) +
  #geom_histogram(alpha=.5, position = "identity") +
  theme_minimal() +
  xlim(c(100, 900)) +
  #scale_fill_manual(values = c("-1" = "black", "0" = "grey",
  scale_color_manual("Region",
                     values = c("-2" = "lightgreen", "-1" = "darkgreen",
                                "0" = "red",
                                "1" = "blue", "2" = "lightblue"),
                     labels = c("precrit-2", "precrit-1", "crit",
                                "spillover1", "spillover2")) + 
  ggtitle("Spread of RTs by Region")
```



## Models 


### Intercept (baseline) LM model

Following the logic of Brouwer et al. (2020), the analysis starts with a baseline model of the variance in the signal. 
The baseline model assumes that no factors were manipulated and that all trials belong to the same condition. 
Therefore, the resulting estimate for each trial in a given region will simply be the average over all trials (and conditions; equivalent to an intercept-only model) in that region. 

Unlike in rERP modeling, different models do not need to be fitted for different electrodes, but like in rERP modeling, different models will be fitted for each subject ($N=40$) and time point (i.e., region, $N=4$: pre-critical region, critical region, post-critical region 1, post-critical region 2), yielding a total of $40 * 4 = 160$ models.

However, here we first start with an even simpler baseline model averaging over subjects.

```{r, results = FALSE}
# Model formula
m.formula <- as.formula(paste0(DV, " ~ 1"))

# Run model(s) and generate plots 
modelAndPlot("Intercept",
             m.formula,
             df,
             m.type,
             regions,
             DV, 
             y.unit, 
             y.range, 
             y.range.res)
```



### Models with a single predictor 

#### Intercept + Association LM

```{r, results = FALSE}
m.formula <- as.formula(paste0(DV, " ~ 1 + Association"))
modelAndPlot(
  "Intercept+Association",
  m.formula,
  df,
  m.type,
  regions,
  DV,
  y.unit,
  y.range,
  y.range.res
)
```


#### Intercept + Coherence LM

```{r, results = FALSE}
m.formula <- as.formula(paste0(DV, " ~ 1 + Coherence"))
modelAndPlot(
  "Intercept+Coherence",
  m.formula,
  df, m.type, regions, DV,
  y.unit, y.range, y.range.res
)
```


#### Intercept + PrecritRT 1 LM

```{r, results = FALSE}
m.formula <- as.formula(paste0(DV, " ~ 1 + ", paste0(precrit1)))
modelAndPlot(
  "Intercept+PrecritRT1",
  m.formula,
  df, m.type, regions, DV,
  y.unit, y.range, y.range.res
)
```



### Models with more than 1 predictor

#### Intercept + Association + Coherence + Precrit_RT1

```{r, results = FALSE}
m.formula <- as.formula(paste0(DV, " ~ 1 + Association + Coherence + ", precrit1))
modelAndPlot(
  "Intercept+Association+Coherence+Precrit1",
  m.formula,
  df, m.type, regions, DV,
  y.unit, y.range, y.range.res
)
```


### Model with interaction term

#### Intercept + Association + Coherence + Association : Coherence + RT_Precrit1

```{r, results = FALSE}
m.formula <- as.formula(paste0(
  DV,
  " ~ Association + Coherence + Association:Coherence + ",
  precrit1
))

modelAndPlot(
  "Intercept+Association+Coherence+Association:Coherence+Precrit1",
  m.formula,
  df, m.type, regions, DV,
  y.unit, y.range, y.range.res
)
```


### lmer 

#### Intercept + Association + Coherence + Precrit1 + Association:Coherence, Random Intercepts 

```{r, results = FALSE}
m.formula <- as.formula(paste0(DV, " ~ 1 + Association + Coherence + Association:Coherence + ",
                               precrit1,
                              " + (1 | Item) + (1 | Subject)"))
modelAndPlot(
  "Intercept+Association+Coherence+Association:Coherence+Precrit1",
  m.formula,
  df, "lmer", regions, DV,
  y.unit, y.range, y.range.res
)
```



### Setting predictors to zero

```{r, fig.height = 7, fig.width = 14}
m.formula <- as.formula(paste0(
  DV,
  " ~ Association + Coherence + Association:Coherence + ",
  precrit1
))

m <- runModels(df,
               "lm.by.subj",
               m.formula,
               regions)

# Set predictors to zero: start with all zero except the intercept,
# then add in more predictors

# Plot only the intercept
m$InterceptOnly <- m$`Coef_(Intercept)`
plotSPR(m, "InterceptOnly", "Estimates: Intercept only", y.unit, y.range)

# Add association only
m$InterceptAssoc <- m$`Coef_(Intercept)` + m$Association*m$Coef_Association 
plotSPR(m, "InterceptAssoc", "Estimates: Intercept+Assoc", y.unit, y.range)

# Add coherence only
m$InterceptCoherence <- m$`Coef_(Intercept)` + m$Coherence*m$Coef_Coherence
plotSPR(m, "InterceptCoherence", "Estimates: Intercept+Coherence", y.unit, y.range)

# Add precritical only 
m$InterceptPrecrit1 <- m$`Coef_(Intercept)` + m$logRT_precrit1*m$Coef_logRT_precrit1
plotSPR(m, "InterceptPrecrit1", "Estimates: Intercept+Precrit1", y.unit, y.range)

# Add everything except association 
m$InterceptCoherencePrecrit1Interaction <- m$`Coef_(Intercept)` + m$Association*0 + m$Coherence*m$Coef_Coherence + m$logRT_precrit1*m$Coef_logRT_precrit1 + m$Association*m$Coherence*m$`Coef_Association:Coherence`
p1 <- plotSPR(m, "InterceptCoherencePrecrit1Interaction", "Estimates: Full model without Association",
        y.unit, y.range)

# Add everything except coherence 
m$InterceptAssocPrecrit1Interaction <- m$`Coef_(Intercept)` + m$Association*m$Coef_Association + m$Coherence*0 + m$logRT_precrit1*m$Coef_logRT_precrit1 + m$Association*m$Coherence*m$`Coef_Association:Coherence`
p2 <- plotSPR(m, "InterceptAssocPrecrit1Interaction", "Estimates: Full model without Coherence",
        y.unit, y.range)

# Add everything except precritical 1 
m$InterceptAssocCoherenceInteraction <- m$`Coef_(Intercept)` + m$Association*m$Coef_Association + m$Coherence*m$Coef_Coherence + m$logRT_precrit1*0 + m$Association*m$Coherence*m$`Coef_Association:Coherence`
p3 <- plotSPR(m, "InterceptAssocCoherenceInteraction", "Estimates: Full model without Precrit1",
        y.unit, y.range)


# Add everything except the interaction term 
m$InterceptAssocCoherencePrecrit1 <- m$`Coef_(Intercept)` + m$Association*m$Coef_Association + m$Coherence*m$Coef_Coherence + m$logRT_precrit1*m$Coef_logRT_precrit1 + m$Association*m$Coherence*0
p4 <- plotSPR(m, "InterceptAssocCoherencePrecrit1", "Estimates: Full model without the interaction term",
        y.unit, y.range)

# Sanity check: full model 
m$Full <- m$`Coef_(Intercept)` + m$Association*m$Coef_Association  + m$Coherence*m$Coef_Coherence + m$logRT_precrit1*m$Coef_logRT_precrit1 +  m$Association*m$Coherence*m$`Coef_Association:Coherence`
p5 <- plotSPR(m, "Full", "Estimates: Full model", y.unit, y.range)

# Observed data 
p.observed <- plotSPR(m, DV, "Observed RTs", y.unit, y.range)

grid.arrange(p1, p2, p3, 
             p4, p5, p.observed,
             nrow=2,
             top = textGrob("Setting individual predictors to zero", gp=gpar(fontsize=14, fontface=2)))
```
