---
title: "Inference SPR results: sampling"
author: "Miriam Schulz"
date: "9/15/2022"
output:
  html_document:
    toc: true
    number_sections: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE)
```


# About

This script randomly samples a subset of participants from the results to provide an idea of the variation in the data.


# Preliminaries

```{r, include = FALSE}
# Workspace, libraries and functions
rm(list = ls())        # clear workspace (optional)
library(tidyverse)
library(grid)
library(gridExtra)     # for grid.arrange
library(sciplot)       # for se function
source("functions.R")  # custom functions 
```

Read in and preprocess data:

```{r}
# Read in preprocessed results file:
reads <- read.csv("results_reads.csv", header = TRUE)

# Remove fillers and practice trials
reads <- removeFillersPractice(reads)

# Keep only regions of interest
reads <- removeRegions(reads, keep.regions = -1:2)

# Add log RTs
reads$logRT <- log(reads$RT)

# Convert variables to numeric / factor:
reads[, c(
  "Subject",
  "IPhash",
  "Cond",
  "Block",
  "Order",
  "Plausible",
  "Region",
  "PlausibilityRating"
)] <-
  lapply(reads[, c(
    "Subject",
    "IPhash",
    "Cond",
    "Block",
    "Order",
    "Plausible",
    "Region",
    "PlausibilityRating"
  )], as.factor)

# Inspect:
str(reads)
```

Exclude data.

Criteria:

- outlier SPR reading times
- outlier context sentence reading times
- incorrect plausibility ratings

```{r}
# Check the original range of the RTs
range(reads$RT)

reads <- removeOutliersSPR(reads,
                        method = "sd",
                        sd.value = 4,
                        entire.trial = TRUE,
                        regions = -1:2)
# Remove outliers: context sentence reading times
reads <- removeOutliersContext(reads, min.rt=500, max.rt=30000)

# Remove trials with incorrect plausibility rating
reads <- removeIncorrect(reads)

# Check the minimum remaining observations per condition:
range(xtabs( ~ Cond + Subject + Region, data = reads))
```

Define a plotting function:

```{r}
# Define a line plot function
line.plot <- function(df) {
  if (log.RTs == TRUE) {
    y.axis.range <- c(5.45, 5.85)
  } else {
    y.axis.range <- c(225, 360)
  }
  p <- df %>%
    ggplot(aes(x = Region, y = Mean, color=Cond, group=Cond)) + 
      geom_point(size=2.5, shape="cross") + 
      geom_line(linewidth=0.5) +
      geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE),
                    width=0.1, linewidth=0.3) +
      geom_vline(xintercept=0, linetype="dashed") +
      #geom_smooth(method = 'lm', se=TRUE, alpha = .2, aes(fill=Cond)) +
      scale_color_manual("Condition",  # Legend title
                         values=c("cornflowerblue", "chartreuse3",
                                  "tomato2", "darkgoldenrod1")) +
      ylim(y.axis.range) +
      theme_minimal() + 
    scale_x_discrete(labels=c("-1" = "Pre-critical", 
                                         "0" = "Critical",
                                         "1" = "Spillover", 
                                         "2" = "Post-spillover"))
  p
}
```


# Sampling

Sample an equal number of participants per list (with replacement across, but not within, individual samples).

```{r fig.height = 10, fig.width = 15}
# Set random seed (try different values)
seed <- 42
set.seed(seed)

# Set parameters and initialize
Nsamples.per.list <- 3
Niterations <- 8
means.cond.list <- list()
plot.list <- list()
log.RTs = TRUE  # if TRUE, use log RTs, else raw RTs

# Sample
for (i in 1:Niterations) {
  
  # Randomly sample from each list.
  # Subjects 1-5 are list 1, order1; subjects 21-25 are list 1, order 2, etc.
  participants.sample <- c(sample(c(1:5, 21:25), Nsamples.per.list),
                           sample(c(6:10, 26:30), Nsamples.per.list),
                           sample(c(11:15, 31:35), Nsamples.per.list),
                           sample(c(16:20, 36:40), Nsamples.per.list))
  reads.sample <- filter(reads, Subject %in% participants.sample)
  
  # Aggregate sampled data 
  if (log.RTs == TRUE) {
    means.cond.sample <-
      means.cond.sample <- aggregMeans(reads.sample,
                                       as.formula(logRT ~ Cond+Region),
                                       n=Nsamples.per.list*4)
  } else {
      means.cond.sample <- aggregMeans(reads.sample,
                                       as.formula(RT ~ Cond+Region),
                                       n=Nsamples.per.list*4)
  }
  means.cond.sample$Iteration <- i  # annotate iteration
  means.cond.list[[i]] <- as.data.frame(means.cond.sample)
  
  # Generate and store plot for current sample 
  p.sample <- line.plot(means.cond.sample)
  plot.list[[i]] <- p.sample
}

# Save or print plots
filename <- paste0("./plots/sample_",
                   Nsamples.per.list*4,
                   "subj_", Niterations,
                   "reps_seed", seed, "_balanced_lists", 
                   ifelse(log.RTs == TRUE, "_log", ""),
                   ".pdf")
#png(file=filename, width = 1200, height = 800)  # good size for 12 samples
pdf(file=filename, width = 8.6, height = 11)   # good size for 8 samples
p <- grid.arrange(grobs = plot.list, ncol=2,
                  top = textGrob(""),
                  gp=gpar(fontsize=4, fontface=2))

                  # Add title
                  # top = textGrob(paste0("Mean ", 
                  #                      ifelse(log.RTs == TRUE, "log ", ""),
                  #                      " RTs for subsets of ",
                  #                      Nsamples.per.list*4,
                  #                      " randomly sampled subjects (",
                  #                      Nsamples.per.list, 
                  #                      " per list)"),
                  #                gp=gpar(fontsize=20, fontface=2)))
dev.off()
plot(p)
```


# Numeric inspection 

Calculate standard deviations of the different samples as a numeric indicator of the variability in the data.

```{r eval = FALSE}
# Turn the list of dataframes for each iteration into a single dataframe
means.cond.df <- bind_rows(means.cond.list)
aggMeansSE(means.cond.df, c("MeanRT", "Cond", "Region"))
```

